% ************************************************
% The Rabbit and Fox growth model
% in RPN code for HP 41C series
%
% (cl) 2013 by Arno Jacobs
%
% An free adaptation of the HP-25 program
%
%

\documentclass[a4paper, landscape]{article}
\usepackage[landscape]{geometry}

\input{../RPN_HP_Base_defs}

% define the title
\author{A.J.M. Jacobs}
\title{The Rabbit and Fox model for the HP-41C family}

\begin{document}
\maketitle
\tableofcontents

\clearpage
\section{The Rabbit and Fox model HP-41C code}

\tStarttabular
\tONfo           {Put the calculator [ON]}
\tPRGMfo         {Enter program mode}
\tGTOendfo       {Set program counter @ end of code} 
\\
\tLBLafo         {FR}      {Start position Rabbit and Fox model}  
\tFIXfo          {5}       {Set the display format to 0.00000}  
\tRCLfo          {02}      {Read amount of rabbits @ time is t}   
\tENTERfo                  {} 
\tENTERfo                  {} 
\tRCLfo          {03}      {Read amount of foxes @ time is t}   
\tRCLfo          {00}      {Encounter factor}   
\tOPmultiplyfo             {}   
\tOPmultiplyfo             {Number of meetings}   
\tSTOfo          {04}      {Store}   
\tXswapYfo                 {}   
\tNBRfo          {2}       {}   
\tOPmultiplyfo             {Calculate change rabbits} 
\tOPminusfo                {} 
\tRCLfo          {01}      {} 
\tOPmultiplyfo             {}
\tSTOPLUSfo      {02}      {Save new number of rabbits} 
\tRCLfo          {04}      {} 
\tRCLfo          {03}      {} 

\tNewPage

\tOPminusfo                {Calculate change foxes} 
\tRCLfo          {01}      {} 
\tOPmultiplyfo             {} 
\tSTOPLUSfo      {03}      {Save new number of foxes} 
\tRCLfo          {03}      {} 
\tEEXfo          {5}       {} 
\tOPdividefo               {Divide number of foxes by 100000} 
\tRCLfo          {02}      {} 
\tXEQfnfo        {INT}     {Add rounded number of rabbits}  
\tOPplusfo                 {Form: rrr.00fff}  
\tRTNfo                    {Return}  
\\
\tGTOendfo                 {END RPN coding} 
\tPRGMfo                   {Leave program mode}
\tASNfo   {FR} {\Rdownfoc} {Assign "FR" to \Rdownfoc } 
\tUSERfo                   {Set USER mode} 
\end{tabular}

\clearpage
\rm
\section{How to use the Fox and Rabbit program}

\bigskip
\it
This model is bases on a publication in a HP Journal (november 1975). It was an example for a HP-25 programmable calculator.
\rm

\bigskip
\noindent
In a fox and rabbit population model the growth functions can be described as a pair of nonlinear differential equations like:

$$ {dR \over dt} = 2 R - \alpha R F $$

$$ {dF \over dt} = -F + \alpha R F $$

\bigskip
\noindent
Where R is the number of rabbits and F the number of foxes. 
The $\alpha$ factor describes the effect of a rabbit meeting a fox.
And t is for time.

\bigskip
With Euler: $ X_{n+1} = X_{n} + h F(X_{n}) $

$$ { R_{n+1} } = { R_{n} + h (2 R_{n} - \alpha R_{n} F_{n}) } $$

$$ { F_{n+1} } = { F_{n} + h ( - F_{n} - \alpha R_{n} F_{n}) } $$


\bigskip
\noindent
The RPN coded Fox and Rabbit program with Label "FR" has been assigned to key 
\Rdownfoc. 
The HP-41C calculator has been set in USER-mode.
\\
Enter and store the initial values.

\bigskip
\bigskip
\bigskip


\noindent
Example:\\
\begin{tabular}{ l l }
 $\alpha = 0.01 $   &  $h = 0.02 $ \\
 $R_0 = 300 $       &  $F_0 = 150 $ \\
\end{tabular} 

\clearpage

\bigskip
\noindent
Keystrokes:\\
\tt
\begin{tabular}{ l l }

 .01 & \STOfo {00} \\
 .02 & \STOfo {01} \\
 300 & \STOfo {02} \\
 150 & \STOfo {03} \\

\end{tabular} 
\rm

\bigskip
\bigskip

\noindent
The program will calculate step by step. Initial values are at time is 0. 
The fist run will result in values at time is 1.

\bigskip
Now run, press key \Rdownfoc

\bigskip
Result:\\
\tt \indent 303.00156 \\ \rm

\noindent
The number before the decimal point is the number of rabbits.
The number after the decimal point is the number of foxes.\\
So, at time is 1 the number of rabbits are 303 and foxes are 156.

\bigskip
Now run again (step 2), press key \Rdownfoc

\bigskip
Result:\\
\tt \indent 305.00162 \\ \rm

At time is 2 the number of rabbits are 305 and foxes are 162.\\
\indent Etc. etc. etc.

\bigskip

\end{document}
